Leadership across cultures and contexts: an exploratory analysis

### 2020-05-09 leadership_across_cultures_contexts.Rmd
Minor edit.

Relationships between leadership and high status, including causality and dependence, are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019]. Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in small-scale warfare or social-political movements). @cheng_dominance_2019 and @cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways. Thus, according to @cheng_why_2020, although all leaders are not necessarily high status, especially across nonhuman species, among humans high social status may be a necessary but insufficient condition for achieving and maintaining positions of leadership. In our view, social context is critical to these discussions [@wiessner_power_2010], which is an idea that is often mentioned but under-developed and under-investigated empirically. Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses. The "low status" male who initiates an inter-group raid on a neighboring group may not be widely respected within his community nor be a candidate for community leadership, however, he might be of exceptionally high status among his peers and high ranking within what @garfield_review_2019 [p. 71] called the "intragenerational social hierarchy."

>>>>

Relationships between leadership and high status, including causality and dependence, are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019]. Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in small-scale warfare or social-political movements). @cheng_dominance_2019 and @cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways. Thus, according to @cheng_why_2020, although all leaders are not necessarily high status individuals, especially across nonhuman species, among humans high social status may most often be a necessary but insufficient condition for achieving and maintaining positions of leadership. In our view, social context is critical to these discussions [@wiessner_power_2010], which is a point often mentioned but under-developed and under-investigated empirically. Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses. The "low status" male who initiates an inter-group raid on a neighboring group may not be widely respected within his community nor be a candidate for community leadership, however, he might be of exceptionally high status among his peers and high ranking within what @garfield_review_2019 [p. 71] called the "intragenerational social hierarchy."
